---
title: "Cluster Analysis"
author: "Réda Es-sakhi"
date: "01/05/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Required packages

```{r, echo=TRUE}
#install.packages("cluster")
#install.packages("factoextra")
#install.packages("cluster.datasets")
#install.packages("xtable")
#install.packages("kableExtra")
#install.packages("knitr")
#install.packages("summarytools")
#install.packages("stargazer")
#install.packages("ade4")
#install.packages("tidymodels")
```

```{r}
x<-c(0,0)
y<-c(6,6)
dist(rbind(x,y),method = "euclidian", diag = T, upper = T)
```
## Excercie 4 minkowski distance 
```{r}
library("ggplot2")
x<-c(0,0)
y<-c(6,6)

MinkowDist=c() # Initialiser à vide la liste
for (p in seq(1,30,.01))
{
MinkowDist=c(MinkowDist,dist(rbind(x, y), method = "minkowski", p = p))     
}

dist_plot <- ggplot(data =data.frame(x = seq(1,30,.01), y=MinkowDist ) , mapping = aes( x=x, y= y))+
  geom_point(size=.1,color="red")+
  xlab("p")+ylab("Minkowski Distance")+
  ggtitle("Minkowski distance wrt p")
 

  dist_plot + annotate("text", x = 3, y = 12, label = "Manhattan, p = 1") + 
    annotate("segment", x = 1, xend = 1, y = 0, yend = 12,
  colour = "blue")+
    annotate("text", x = 4, y = 9, label = "Euclidian, p = 2") + 
    annotate("segment", x = 2, xend = 2, y = 0, yend = 8.5,
  colour = "Green")
  
  
```


## exo5 
```{r}

n = 100
m = log(2500*sqrt(2/13))
s = sqrt(log(26/25))
a <- rlnorm(n,m,s)
mean(a)
sd(a)
plot(a)
```

## exo 6 
```{r}
x = c(22, 34, 1, 12, 25, 56, 7)
y = c(2,64,12,2,22,5,8)
```
### ex6 The rank for each vector

```{r}
r_x = rank(x)
r_y = rank(y)
r_x
r_y
```

### ex6 The rank for each vector
```{r}
d=r_x-r_y
d
```


```{r}
alt_cor = 1 - 6*sum(d^2)/(7*(7^2-1))
alt_cor
```


```{r}
spear_cor = cor(x,y, method=c("spearman"))
spear_cor
```
## exo 7 
x=(22,34,1,12,25,56,7)
y=(2,64,12,2,22,5,8)
1- liste les pairs (faire un tableau) 
2 - nombre de pairs 
Pair of coordinate (2 parmi n ) soit (n*(n-1))/2
```{r}
7*6/2
```
3- for each pair compute the sign (faire un tableau)
```{r}
tau=0
for (i in 1:7)
{  
tau=tau+sign(x -x[i])%*%sign(y -y[i])
}
tau=tau/(7*6)
tau
```
```{r}
cor(x,y,method = "kendall")
```
## exo 8
Age in (years) and height in (cm) of fout people
```{r}
table3 =c(
  35,190,
  40,190,
  35,160,
  40,160
)
table3=data.frame(matrix(table3, nrow=4,byrow=T))
row.names(table3)=c("A","B","C","D")
names(table3)=c("Age", "Height")
table3
```
```{r}
plot(Height~Age, 
     main = "Plot of height (in centimeters) versus age for four people",
     xlab = "Age (YEARS)", ylab = "Height (CM)",
     col="lightblue", pch=19, cex=2,data=table3)
text(Height~Age, labels=rownames(table3), data=table3, cex=0.9, font=2)
```
### Conversion de Height en Feet
```{r}
yCm <-table3$Height
yCm
yFeet <- yCm/30.48
yFeet<-round(yFeet,2)
yFeet

```
### faire un tableau avec kable
```{r}
table4 <- data.frame(table3$Age, yFeet)
row.names(table4)=c("A","B","C","D")
names(table4)=c("Age","Height")
table4
```
### nouveau plot height en Feet
```{r}
plot(Height~Age, 
     main = "Plot of height (in Feet) versus age for four people",
     xlab = "Age (YEARS)", ylab = "Height (ft)",
     col="red", pch=19, cex=2,data=table4)
text(Height~Age, labels=rownames(table4), data=table4, cex=0.9, font=2)
```
On remarque la même distribution sur notre plot. On transforme l'unité de mesure ne change pas la visualisation et la relation entre les deux variables

###
```{r}
xSt <- table3$Age
xSt <- scale(xSt)
xSt
ySt <- table3$Height
ySt <- scale(ySt)
ySt



```
```{r}
(table3$Age - mean(table3$Age))/sd(table3$Age)
```



### Tableau des variables standardisées 
```{r}
table5 <- data.frame(round(xSt,2),round(ySt,2))
row.names(table5) = c("A","B","C","D")
names(table5) = c("Age", "Height")
table5
```
### plot des variables standardisées 
```{r}
plot(Height~Age, 
     main = "Plot of standarized height and age for four people",
     xlab = "Age ST", ylab = "Height ST",
     col="red", pch=19, cex=2,data=table5)
text(Height~Age, labels=rownames(table5), data=table5, cex=0.9, font=2)
```

La même distribution pour les variables standardisées sur notre plot

## exo 9 
Use the data set animals available in the package cluster.
This data set was first used in this textbook KAUFMAN, Leonard et ROUSSEEUW, Peter J. Finding groups in data: an introduction to cluster analysis. John Wiley & Sons, 2009.
Identify the missing measurements.
Explain the way how KAUFMAN and ROUSSEEUW, pp. 296-297 treat the missing measurements.
Compute a distance matrix for the completed data.
Propose a graphical way to represent that distance matrix.
Which group of animals look close?
Change the method of calculating and observe if it has some effect of the graph

### call library and animals dataset
```{r}
library(cluster)
library(stargazer)
library(forcats)
data("animals")
```

### finding missing values
```{r}
head(animals)
is.na(animals)
str(animals)
animalsF<- animals
animalsF[animalsF == "1"] <- 0
animalsF[animalsF == "2"] <- 1
animalsF
animalsF$war <- factor(animalsF$war, levels = c(0,1))
animalsF$fly <- factor(animalsF$fly, levels = c(0,1))
animalsF$ver <- factor(animalsF$ver, levels = c(0,1))
animalsF$end <- factor(animalsF$end, levels = c(0,1))
animalsF$gro <- factor(animalsF$gro, levels = c(0,1))
animalsF$hai <- factor(animalsF$hai, levels = c(0,1))

animalsF
is.na(animalsF)
```
```{r}
animals%>% is.na
```
```{r}
row.names(animals)
```





```{r, include=FALSE}
library(dplyr)
library(tidyverse)
library(tidymodels)
```


```{r}
missing_values <- animalsF %>%
  mutate(row_num = row_number()) %>%
  gather(key = "key", value = "value", -row_num) %>%
  filter(value %>% is.na()) %>%
  #count(row_num, sort = TRUE) %>% 
  select(row_num)

```
```{r}
missing_values
animalsF[missing_values$row_num,]

```
```{r}
animalsFC <- animalsF  # animalsFC animals data completed
animalsFC
animalsFC["lio","end"] <- 1
animalsFC["spi","end"] <- 0
animalsFC["fro","gro"] <- 0
animalsFC["lob","gro"] <- 0
animalsFC["sal","gro"] <- 0
animalsFC%>%is.na()
```

### matrice de distance
```{r}
library(ade4)
df_animalsFC <- as.data.frame(lapply(animalsFC, as.numeric),row.names = c(row.names(animalsFC)))
df_animalsFC
df_animalsFC[df_animalsFC == "1"] <- 0
df_animalsFC[df_animalsFC == "2"] <- 1
df_animalsFC
jaccard_mat <- dist.binary(df_animalsFC, method = 1, diag = T, upper = T)
jaccard_mat
```
### Propose a graphical way to represent that distance matrix
```{r, fig.cap="heatmap for distance matrix - Jaccard"}
jaccard_mat%>%str
heatmap(as.matrix(jaccard_mat),symm = T,keep.dendro = FALSE,Rowv = NA)
```

### simple matching matrix
```{r}
simple_mat <- dist.binary(df_animalsFC, method = 2, diag = T, upper = T)

```
```{r, fig.cap= "Heatmap for distance matrix - simple matching"}
library(factoextra)
fviz_dist(simple_mat)
```

## exo 10

Prove that the distances based on the Simple Matching coefficient and the Jaccard coefficient satisfy A3.
Prove that the distances proposed by Gower and Legendre (1986) do not satisfy A3.
Hint: Proofs and counterexamples have to be adapted from in the paper: Gower, J. C., & Legendre, P. (1986). Metric and Euclidean properties of dissimilarity coefficients. Journal of classification, 3(1), 5-48.

(Inclure les preuves de :
https://arxiv.org/pdf/1612.02696.pdf

https://www.researchgate.net/publication/24056133_Metric_and_Euclidean_properties_of_dissimilarity_coefficients/link/00463532856f248ec4000000/download



## exo 11
The number of possible partition of n objects is given by the Stirling numbers of the second kind


$$
{n\brace k} = {1\over k!} \sum_{i=1}^{k} (-1)^{i}{ n\choose k} (k-i)^{n} 
$$

example function in R 
How many partitions of 3 objects are they in 6 object


```{r}
#install.packages("multicool")
library(multicool)
Stirling2(6,3)

```

## exo 13
```{r}
df=read.csv("https://raw.githubusercontent.com/karkil2205/Cluster-Analysis/master/Hartigandata1.csv")
df%>%head
df%>%summary
```
```{r}
df<-df[1:8,c(3,4,6)]
df

df[3,1]

df[3,1]<-13 # Error in line 3
df[6,1]<-4 # Error at line 6
df[7,3]<-1 # Error at line 7
df

rownames(df)<-c("BB","HR","BR","BS","BC","CB","CC","BH")
colnames(df)<-c("Energy","Protein","Calcium")
df
```


```{r}
km.res<-kmeans(df[1:8,],3,iter.max = 100)
km.res
```


```{r}
k_clus <- km.res$cluster
k_clus

```


```{r}
km.res$centers
km.res$totss
sum((df[1:8,]$Energy-mean(df[1:8,]$Energy))^2)+
  sum((df[1:8,]$Protein-mean(df[1:8,]$Protein))^2)+
  sum((df[1:8,]$Calcium-mean(df[1:8,]$Calcium))^2)
7*var(df[1:8,]$Energy)+7*var(df[1:8,]$Protein)+7*var(df[1:8,]$Calcium)
```


```{r}
df
```


### rearrange with k  3 and the found clustering

```{r}
df_ar <- df
k_clus <- as.data.frame(k_clus)
k_clus
df_ar <- bind_cols(k_clus, df) %>% arrange(k_clus)
df_ar
```
### heatmap with data arranged 
```{r, fig.cap= "heatmap with data arranded according to the found clusters - k = 3"}
df_ar %>% select(-k_clus) %>% dist(diag = T, upper = T ) %>% as.matrix %>%
  heatmap(symm = T,keep.dendro = T, Rowv = NA, )
```

exo 15 

```{r}
k3_cluster <- km.res$centers
k3_cluster %>% as.data.frame() %>% rownames_to_column("cluster_num")
  #add_column()
```


```{r}
k_clus %>% arrange(k_clus)
cluster_nut <- data.frame(c("1","2","3"),c(("HR BC CB"),("BB BR BS"),("CC BH")))
names(cluster_nut) = c("cluster_num","cluster")
k3_cluster_num <-k3_cluster %>% as.data.frame() %>% rownames_to_column("cluster_num")
table15 <-left_join(k3_cluster_num,cluster_nut)
table15 %>% select(c("cluster_num","cluster","Energy","Protein","Calcium"))
```


```{r}
km.res
```


```{r}
km.res$withinss
numb_of_erro <- km.res$withinss
```


```{r}
plot(x = table15$cluster_num , y = numb_of_erro)
```



## exo 18 
```{r}
fviz_nbclust(df, kmeans,method = "silhouette", k.max = 7)
```

```{r}
k4 <- kmeans(df,4,iter.max = 100)
sil_obj <- silhouette(k4$cluster, dist(df))
sil_obj %>% fviz_silhouette(print.summary = T)
```


```{r}

km.res$cluster
slobj <- silhouette(km.res$cluster, dist(df_ar))
slobj %>% fviz_silhouette() 

```



## exo 19
```{r}
library(datasets)
data("USArrests")
USArrests
```

```{r}
fviz_nbclust(USArrests, kmeans,method = "silhouette", k.max = 49)
```

```{r}
km.arrest <- kmeans(USArrests,2,iter.max = 100)
km.arrest
```
```{r}
fviz_cluster(km.arrest, USArrests, ellipse.type = "norm")
```


```{r}
fviz_cluster(km.arrest, USArrests,
palette = c("#2E9FDF", "#00AFBB", "#E7B800"),
ellipse.type = "euclid", # Concentration ellipse
star.plot = TRUE, # Add segments from centroids to items
repel = TRUE, # Avoid label overplotting (slow)
ggtheme = theme_minimal()
)
```
```{r}
silhouette(km.arrest$cluster,dist(USArrests))%>%
fviz_silhouette()
```






