---
title: 'Discrete Choice Models'
author: "Réda Es-sakhi"
date: "09/05/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Origin of the data
For the Discrete Choise Models project, we are using Yogurt data available in Ecdat's package


```{r load library, include = FALSE}
#install.packages("knitr")
#install.packages("kableExtra")
#install.packages("mlogit")
#install.packages("Ecdat")
library(knitr)
library(mlogit)
library(Ecdat)
data("Cracker",package="mlogit")
data("Yogurt",package="Ecdat")
```

```{r head cracker data}
head(Yogurt)
```


# Data preparation 
* Yogurt data is transformed from wide format to long format where each column considered as variable.


```{r , fig.cap= "table 1 : Yogurtlong summary statistics"}
Yogurtlong<-mlogit.data(Yogurt,shape = "long",varying = 2:9,choice = "choice")
library(stargazer)
names(Yogurtlong)[3]<-c("Brand")
stargazer(Yogurtlong, summary=T,type = 'text',title = "table 1 : Yogurtlong summary statistics")


```

## summarising Yogurt data by brand
Here we analyze each brand data. Market Share (MS), Price and Feat. 
```{r summarise yogurt data}
library(dplyr)
MS_tab=summarise(
  group_by(Yogurtlong,Brand),
          mean=mean(choice),
          sd=sd(choice),
            )
Price_tab=summarise(
  group_by(Yogurtlong,Brand),
          mean=mean(price),
          sd=sd(price),
            )
Feat_tab=summarise(
  group_by(Yogurtlong,Brand),
          mean=mean(feat),
          sd=sd(feat),
              )
```


```{r kable yogurt data}
kables(
  list(
    kable(MS_tab[1],digits = 4,caption = "Brand"),
    kable(MS_tab[-1],digits = 4,caption = "Market Shares"),
    kable(Price_tab[-1],digits = 4,caption = "Prices"),
    kable(Feat_tab[-1],digits = 4,caption = "Feature")
  ),
  caption = "Yogurt Data summary statistics"
) #%>% kableExtra::kable_styling(bootstrap_options = "basic", full_width = F)
```
## MNL Model



MNL Models or Multinomial logit model is given by the equation :  
$$
\mathbb{P}_i = \frac{e^{u_i}}{\sum_{k=1}^ne^{u_k}}, \:i=1,\cdots,n
$$
where $u_1,\cdots,u_n$ deterministic or systematic part and $n$ the number of alternatives.  
Applied to our data we have for choice probabilities the following : 
```{r}
beta = 1
log(exp(beta*Yogurtlong$price[1:4])/sum(exp(beta*Yogurtlong$price[1:4])))%*%Yogurtlong$choice[1:4]
```
li_func 
```{r,fig.cap= "Variation of Likelihood"}
li_func <- function(beta){
  somme = 0
  for (i in seq(0.9644,4)) {
    loga <- log(exp(beta*Yogurtlong$price[(i+1):(i+4)])/sum(exp(beta*Yogurtlong$price[(i+1):(i+4)])))%*%Yogurtlong$choice[(i+1):(i+4)]
    somme <- loga + somme
  }
  return(somme)
}
resLi = c()
for (beta in seq(-1,1,0.1)) {
  resLi = c(resLi,li_func(beta))
}
plot(x=seq(-1,1,0.1),y=resLi)
```






```{r}
optimize(li_func, interval = c(-1,1),maximum = TRUE)
```
```{r}
library("mlogit")
m<-mlogit(choice~price+feat,Yogurtlong)
stargazer(m,summary=F,type = 'text')
```




